Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-07 shanghai.tex

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
If $\cs{L} = (\gcs{L}, \mu, F)$ is a character sheaf on $G$ we may define
a character sheaf $\nu(\cs{L}) = (\gcs{L}, \mu, F_{k'})$ on $G_{k'}$ by setting
\[
F_{k'} = F \circ \Frob_G^*(F) \circ \cdots \circ (\Frob_G^{n-1})^*(F).
\]
The commutativity of the diagram (CS.2) for $F_{k'}$ follows from the fact that $\Frob_{G_{k'}} = \Frob_G^n$.
Note that we may also think about the construction of $F_{k'}$ as taking the action of $\Weil{k}$ defined in Remark
\ref{rem:CS-action} and restricting it to $\Weil{k'}$.

\begin{proposition}
The following diagram commutes, where the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}
\CSiso{G} \rar{\nu} \dar{t_G} & \CSiso{G_{k'}} \dar{t_{G_{k'}}} \\
\Hom(G(k), \EEx) \rar{\Nm} & \Hom(G(k'), \EEx).
\end{tikzcd}
\]
\end{proposition}

In the opposite direction, there is a canonical closed immersion of $k$-schemes
$G \hookrightarrow \Res_{k'/k}(G_{k'})$ \cite[\S 7.6]{BLR}.  Pullback yields a functor
\[
\CS(\Res_{k'/k}(G_{k'})) \xrightarrow{\iota} \CS(G).
\]
\begin{proposition} \label{prop:csbe}
The following diagram commutes, where the bottom map is induced by the inclusion $G(k) \hookrightarrow G(k')$:
\[
\begin{tikzcd}
\CSiso{\Res_{k'/k}(G_{k'})} \arrow[two heads]{r}{\iota} \dar{t_{\Res_{k'/k}(G_{k'})}} & \CSiso{G} \dar{t_G} \\
\Hom(G(k'), \EEx) \arrow[two heads]{r} & \Hom(G(k), \EEx).
\end{tikzcd}
\]
\end{proposition}

Now suppose that $G'$ is an abelian group scheme over $k'$, satisfying the conditions of Theorem \ref{thm:snake}.
There is a canonical isomorphism between $G'(k')$ and $(\Res_{k'/k}G')(k)$;
we now describe the corresponding functor from
$\CS(G')$ to $\CS(\Res_{k'/k}G')$.  The base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
decomposes into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
The component corresponding to the identity element yields a natural inclusion 
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'}
\]
of $k'$-schemes.  The definition of base change yields a map of $k$-schemes
\[
(\Res_{k'/k}G')_{k'} \to \Res_{k'/k}G'.
\]
Consecutive pullback along these maps yields a functor
\[
\CS(G') \to \CS(\Res_{k'/k}G').
\]

\begin{proposition}
  Then the following diagram commutes:
\[
\begin{tikzcd}
\CSiso{G'} \dar{t_{G'}} \rar & \CSiso{\Res_{k'/k} G'} \dar{t_{\Res_{k'/k} G'}} \\
 \Hom(G'(k'), \EEx) \rar & \Hom((\Res_{k'/k}G')(k), \EEx)
\end{tikzcd}
\]
\end{proposition}

@@

\section{Functorial properties of the main isomorphism}

We have already seen that the trace of Frobenius defines a canonical isomorphism 
$t_T : \CSiso{\GN{T}} \to \Hom_\text{ad}(T(K),\EEx)$. 
In this section we expand on the canonicity of this group isomorphism 
by showing that it is a natural transformation. We also show

\begin{proposition}
The isomorphism $t_T : \CSiso{\GN{T}} \to \Hom_\text{ad}(T(K),\EEx)$ 
is a natural transformation between the following two contravariant functors 
from the category of algebraic tori over $K$ to the category of abelian groups: 
\begin{enumerate}
\item $T \mapsto \CSiso{\GN{T}}$
\item $T \mapsto \Hom_\text{ad}(T(K),\EEx)$.
\end{enumerate}
\end{proposition}

\Clifton{I think I can prove this proposition without too much difficulty.}
\begin{proof}
\end{proof}

Let $L/K$ be a finite Galois extension and let $K'/K$ be the maximal unramified subextension, with residue field $k'$.  If $T$ is a torus over $K$, write $\TL$ for the N\'eron model of the base change $T_L$.  Then there is a canonical isomorphism
\[
\Res_{k'/k} \Gr(\TL) \cong \Gr(\underline{\Res_{L/K}(T_L)}).
\]
The natural inclusion $T \hookrightarrow \Res_{L/K} T_L$ induces a map of $k$-schemes
\[
\GN{T} \hookrightarrow \Res_{k'/k} \Gr(\TL)
\]
and thus a map on character sheaves.

\begin{proposition}
The following diagram commutes:
\[
\begin{tikzcd}
\CSiso{\Res_{k'/k} \Gr(\TL)} \arrow[two heads]{r} \dar & \CSiso{\GN{T}} \dar\\
\Hom(T(L), \EEx) \arrow[two heads]{r} & \Hom(T(K), \EEx)
\end{tikzcd}
\]
\end{proposition}

We now translate the norm map $T(L) \to T(K)$ to character sheaves.  There is an open immersion
\[
\TT \times_{\OK} \OL \to \TL
\]
induced by the canonical isomorphism of generic fibers \cite[p. 1]{Chai}.  After taking Greenberg transforms,
we obtain a diagram
\[
\begin{tikzcd}
\Gr(\TT) \dar & \Gr(\TT) \times_{\Gr(\OK)} \Gr(\OL) \lar{\alpha} \rar{\beta} \dar & \Gr(\TL) \dar \\
\Spec{k} & \Spec{k'} \lar \rar & \Spec{k'}
\end{tikzcd}
\]
In the case that $L/K$ is unramified, the left hand square is a pullback square and the right hand horizontal maps are isomorphisms.  In general, we may define a functor
\begin{align*}
\CS(\GN{T}) &\to \CS(\Gr(\TL)) \\
\cs{L} &\mapsto \beta_* \alpha^* \cs{L}.
\end{align*}

\begin{proposition}
The following diagram commutes:
\[
\begin{tikzcd}
\CSiso{\GN{T}} \rar{\beta_* \circ \alpha^*} \dar & \CSiso{\Gr(\TL)} \dar \\
\Hom(T(K), \EEx) \rar{\Nm} & \Hom(T(L), \EEx)
\end{tikzcd}
\]
\end{proposition}

@@

>>>>

@@

\section{The transfer principle}

\David{We should add a discussion of why we care about the transfer principle.}
We recall the notion of $N$-congruent tori from \cite[\S 2]{CY}.  Suppose $T$ and $T'$
are tori over non-Archimedian local fields $K$ and $K'$, splitting over $L$ and $L'$ respectively.

\begin{definition}
 We say that $T$ and $T'$ are \emph{$N$-congruent} if there are isomorphisms
 \begin{align*}
  \alpha : \OL/\pi_K^N \OL &\to \OO{L'}/\pi_{K'}^N \OO{L'} \\
  \beta : \Gal(L/K) &\to \Gal(L'/K') \\
  \phi : X^*(T) &\to X^*(T')
 \end{align*}
 satisfying the following conditions:
 \begin{enumerate}
  \item $\alpha$ induces an isomorphism $\OK/\pi_K^N \OK \to \OO{K'}/\pi_{K'}^N \OO{K'}$,
  \item $\alpha$ is $\Gal(L/K)$-equivariant relative to $\beta$: $\alpha(\gamma x) = \beta(\gamma) \alpha(x)$
  for $\gamma \in \Gal(L/K)$ and $x \in \OL/\pi_K^N \OL$,
  \item $\phi$ is $\Gal(L/K)$-equivariant relative to $\beta$.
 \end{enumerate}
\end{definition}

One of the main results of \cite{CY} gives an isomorphism of group schemes between
$\TT_d$ and $\TT'_d$ assuming that $T$ and $T'$ are sufficiently congruent.
They define a quantity $h$ as the smallest integer so that $\pi^h$ lies in the
Jacobian ideal associated to a natural embedding of $T$ into an induced torus \cite[\S 8.1]{CY}.
They then prove the following theorem:

\begin{theorem}[{\cite[Thm. 8.5]{CY}}]
 Suppose that $T$ and $T'$ are $N$-congruent for $N > 3h$.  Then there is a canonical isomorphism
 \[
  \TT_{N-3h} \to \TT'_{N-3h}
 \]
 determined by $\alpha, \beta$ and $\phi$.
\end{theorem}

As an immediate consequence of this theorem and Theorem \ref{thm:application} we have the following corollary:

\begin{corollary}
 Suppose that $T$ and $T'$ are $N$-congruent for $N > 3h$.  Then there is a canonical equivalence of categories
 \[
  \CS(\GN{T}_{N-3h}) \to \CS(\GN{T}'_{N-3h})
 \]
 determined by $\alpha, \beta$ and $\phi$ inducing the isomorphism
 \[
  \Hom(T(K) / T(K)_{N-3h}, \EEx) \to \Hom(T'(K') / T'(K')_{N-3h}, \EEx)
 \]
\end{corollary}

@@

\bib{CY}{article}{
   author={Chai, C.L.},
   author={Yu, J.-K.},
   title={Congruences of N\'eron models for tori and the Artin conductor},
   journal={Ann. of Math.},
   volume={154},
   number={2},
   date={2001},
   pages={347-382}}
